A Survey of Sterile Neutrinos

### 2015-04-15 p3_505/paper/paper3.tex
done?

\documentclass[aps,prd,twocolumn,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{listings}
\usepackage[hidelinks,hyperfootnotes=false,bookmarks=false]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{verbatim}
\usepackage{float}
\begin{document}
\title{A Survey of Sterile Neutrinos}
\author{Douglas Davis}
\author{Matthew Epland}
\author{Justin Raybern}
\author{Pingchuan Zhao}
\affiliation{Department of Physics, Duke University, Durham, NC 27707, USA}
\date{\today}
\begin{abstract}
Neutrinos are light, weakly interacting, left-handed leptons in the Standard Model of particle physics. The discovery of neutrino oscillations (suggesting neutrinos actually have mass) has led to a fruitful time in neutrino research. Oscillation experiments measure the oscillation parameters $\Delta m^2$ and $\sin^2 2\theta$, which are the mass splittings of neutrino mass eigenstates and the mixing angles of a unitary transformaton matrix that governs the oscillation probabilities. Experiments have measured these parameters, and some results suggest the existence of the expected three weakly interacting neutrinos, but also the possibility of more neutrinos which are not weakly interacting, and are therefore given the name ``sterile'' neutrino. We summarize the theory and experimental evidence in the sterile neutrino sector.
\end{abstract}\maketitle
\section{Theory}
\label{sec:theory}
\subsection{Standard Model and Beyond Standard Moden Neutrinos}
A neutrino is an electrically neutral, weakly interacting elementary particle with half integer spin. There are three generations of leptons and their corresponding neutrino: electron \& electron-neutrino, muon \& muon-neutrino, and tau \& tau-neutrino. Neutrinos were first postulated as products in beta decay. From the experiment of radioactive decay, the outgoing electron energy varies; therefore it is not a 2-body decay, but actually a 3-body decay. Pauli first called that small, neutral charge particle neutrino.

@@

\subsection{Theory Justin}

@@

where $\nu$ and $N$ are the left- and right-handed neutrino components, respectively. The Dirac mass is given by $m_D$ and the Majorana mass is given my $M$. It is easy to show that the eigenvalues of this matrix, if $m_D \ll M$, are $m_D^2/M$ and $M$ with the approximate eigenvectors associated with the left- and right- handed neutrinos. This means that we have the normal neutrino eigenstate and a sterile neutrino eigenstate.

@@

\section{Conclusions}
xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx

\begin{thebibliography}{9}
\bibitem{ping1}
  T2K Experiment -- \url{http://t2k-experiment.org/neutrinos/in-the-standard-model/}

@@

\bibitem{justin1}
Biswajoy Brahmachari, A see-saw model of sterile neutrino, \emph{Physics Letters B}, {\bf 461} 3, 26 August 1999, Pages 243-247.
\bibitem{LSND}
  A.~Aguilar-Arevalo \emph{et al.} Evidence for neutrino oscillations from the observation of anti-neutrino(electron) appearance in a anti-neutrino(muon) beam. \emph{Phys. Rev. D.}, {\bf 64} 112007, 2001.
\bibitem{mini1}
  A.~Aguilar-Arevalo \emph{et al.} Event Excess in the MiniBooNE Search for $\nu_\mu \rightarrow \nu_e$ Oscillations. \emph{Phys. Rev. Lett.}, {\bf 105} 181801, 2010.
\bibitem{mini2}
  A.~Aguilar-Arevalo \emph{et al.} Improved Search $\nu_\mu \rightarrow \nu_e$ Oscillations in the MiniBooNE Experiment. \emph{Phys. Rev. Lett.}, {\bf 110} 161801, 2013.
\bibitem{reactor_anom1}
  G. Mention, M. Fechner, Th. Lasserre, Th.A. Mueller, D. Lhuillier, \emph{et al.} The Reactor Antineutrino Anomaly. \emph{Phys. Rev. D.}, {\bf 83} 073006, 2011.
\bibitem{reactor_anom2}
  Th. A. Mueller, D. Lhuillier, M. Fallot, A. Letourneau, S. Cormon, \emph{et al.} Improved Predictions of Reactor Antineutrino Spectra. \emph{Phys. Rev. C.}, {\bf 83} 054615, 2011.
\bibitem{gallex1}
  P.~Anselmann~\emph{et al.} First results from the Cr-51 neutrino source experiment with the GALLEX detector. \emph{Phys. Lett. B.}, {\bf 342} 440-450, 1995
\bibitem{gallex2}
  W.~Hampel~\emph{et al.} Final results of the Cr-51 neutrino source experiments in GALLEX. \emph{Phys. Lett. B.}, {\bf 420} 114-126, 1998.
\bibitem{sage1}
  Dzh.~N.~Abdurashitov, V.N.~Gavrin, S.V. Girin, V.V. Gorbachev, Tatiana V. Ibragimova, \emph{et al.} The Russian-American gallium experiment (SAGE) Cr neutrino source measurement. \emph{Phys. Rev. Lett.}, {\bf 77} 4708-4711, 1996.
\bibitem{sage2}
  J.~N.~Abdurashitov~\emph{et al.} Measurement of the response of the Russian-American gallium experiment to neutrinos from a Cr-51 source. \emph{Phys. Rev. C.}, {\bf 59} 2246-2263, 1999.
\bibitem{MINOS}
  Alexandre B. Sousa (for the MINOS and MINOS+ Collaborations) First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}
\bibitem{SuperK}
  K. Abe~\emph{et al.} (Super-Kamiokande Collaboration) Limits on sterile neutrino mixing using atmospheric neutrinos in Super-Kamiokande. \emph{Phys. Rev. D}, {\bf 91}, 052019, 2015, \url{http://dx.doi.org/10.1103/PhysRevD.91.052019}
\bibitem{DayaBay}
  F. P. An~\emph{et al.} (Daya Bay Collaboration) Search for a Light Sterile Neutrino at Daya Bay. \emph{Phys. Rev. Lett.} {\bf 113}, 141802, 2014, \url{http://dx.doi.org/10.1103/PhysRevLett.113.141802}

>>>>

\documentclass[aps,prd,twocolumn,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{listings}
\usepackage[hidelinks,hyperfootnotes=false,bookmarks=false]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{verbatim}
\usepackage{float}
\begin{document}
\title{A Survey of Sterile Neutrinos}
\author{Douglas Davis}
\author{Matthew Epland}
\author{Justin Raybern}
\author{Pingchuan Zhao}
\affiliation{Department of Physics, Duke University, Durham, NC 27707, USA}
\date{\today}
\begin{abstract}
Neutrinos are light, weakly interacting, left-handed leptons in the Standard Model of particle physics. The discovery of neutrino oscillations (suggesting neutrinos actually have mass) has led to a fruitful time in neutrino research. Oscillation experiments measure the oscillation parameters $\Delta m^2$ and $\sin^2 2\theta$, which are the mass splittings of neutrino mass eigenstates and the mixing angles of a unitary transformaton matrix that governs the oscillation probabilities. Experiments have measured these parameters, and some results suggest the existence of the expected three weakly interacting neutrinos, but also the possibility of more neutrinos which are not weakly interacting, and are therefore given the name ``sterile'' neutrino. We summarize the theory and experimental evidence in the sterile neutrino sector.
\end{abstract}\maketitle
\section{Introduction}
The Standard Model of Particle Physics continues to succeed wonderfully in making predictions in nature.  The discovery of the Higg's Boson is just one example of the predictive power of the model.  There are places in nature, however, where the Standard Model fails - the fact that neutrinos have mass, which is confirmed by flavor oscillations, is completely outside the model.  The smallness of the neutrino mass cannot be explained definitively by theory either.  And even more interestingly, neutrinos seem to only come as left-handed particles.

In this work we lay out some of basic Standard Model theory concerning neutrinos.  We then describe oscillations and the significance of that behavior.  We then explain sterile neutrinos - neutrinos that do not interact via the weak force, and how the see-saw model of neutrino mass can simultaneously explain light, left-handed neutrinos.  We discuss some of the contested experimental evidence for sterile neutrinos and future experimental work in the field.
\section{Theory}
\label{sec:theory}
\subsection{Standard Model, and Beyond, Neutrinos}
A neutrino is an electrically neutral, weakly interacting elementary particle with half integer spin. There are three generations of leptons and their corresponding neutrino: electron \& electron-neutrino, muon \& muon-neutrino, and tau \& tau-neutrino. Neutrinos were first postulated as products in beta decay. From the experiment of radioactive decay, the outgoing electron energy varies; therefore it is not a 2-body decay, but actually a 3-body decay. Pauli first called that small, neutral charge particle neutrino.

@@

Oscillation and survival probabilities for $\nu_\mu\rightarrow\nu_e$ are shown in Figure~\ref{fig:probs}.
\begin{figure}[H]
  \centering
  \includegraphics[width=1\linewidth]{../figures/oscillation_prob.pdf}
  \caption{Oscillation and survival probabilities for the two neutrino state with an initial $\nu_{\mu}$. The values for $\Delta m^2$ and $\sin^22\theta$ ($2.41\times 10^{-3}$ eV$^2$ and 0.97, respectively) are taken from 2013 MINOS result (PRL, 110, 251801)}
  \label{fig:probs}
\end{figure}
\subsection{See-Saw Neutrinos}

@@

where $\nu$ and $N$ are the left- and right-handed neutrino components, respectively. The Dirac mass is given by $m_D$ and the Majorana mass is given my $M$. It is easy to show that the eigenvalues of this matrix, if $m_D \ll M$, are $m_D^2/M$ and $M$ with the approximate eigenvectors associated with the left- and right- handed neutrinos. This means that we have the normal neutrino eigenstate and a sterile neutrino eigenstate. One can see that as $M$ increases or decreases $m_D^2/M$ does the opposite -- thus the name ``see-saw.'' See Figure~\ref{fig:seesaw}.
\begin{figure}[H]
  \centering
  \includegraphics[width=.84\linewidth]{../figures/seesaw.pdf}
  \caption{An illustration of the see-saw mechanism with left handed and right handed neutrino masses}
  \label{fig:seesaw}
\end{figure}

There are different see-saw models associated with the introduction of isospin singlets, scalar triplets, and fermion triplets for types I, II, and III respectively.  The Type I See-Saw model is by far the most popular with theorists.  It suffers from requiring an incredibly large majorana mass around $10^{13}$ GeV making it untestable in current particle accelerators~\cite{justin2}. Type II models also rely on quite large Majorana masses. Type III models have the best chance at experimental verification but are the least phenomenologically satisfying and are thus the least popular.

@@

\section{Conclusions}
The discovery of heavy, right-handed sterile neutrinos would help explain some of the mystery surrounding the nature of neutrinos.  While there is some evidence for sterile neutrinos from LSND and MiniBooNE, it is very hotly contested evidence.  Further work remains to be done before the answer is clear.  Neutrino specific experiments like MINOS+, Super-K, and Daya Bay are continuing the search for evidence for sterile neutrinos.  As the amount of data continues to increase, the anomalies will certainly be confirmed or refuted very soon.

\begin{thebibliography}{9}
\bibitem{ping1}
  T2K Experiment -- \url{http://t2k-experiment.org/neutrinos/in-the-standard-model/}
\bibitem{ping2}
Solar neutrinos. Rep. Prog. Phys. 64 (2001) 97–146
\bibitem{ping3}
Daya Bay Experiment: \url{https://www.bnl.gov/science/dayabay.php}
\bibitem{ping4}
Neutrinoless double beta decay and CP violation Physics Letters B Volume 157, Issues 5–6, 25 July 1985, Pages 442–446
\bibitem{ping5}
Neutrinoless Double Beta Decay Experiments. arXiv:1408.2455v1 [hep-ex] 11 Aug 2014

@@

\bibitem{justin1}
Biswajoy Brahmachari, A see-saw model of sterile neutrino, \emph{Phys. Lett. B,}, {\bf 461} 3, 26 August 1999, Pages 243-247.
\bibitem{justin2}
W. Grimus, L. Lavoura, B. Radovcic, Type II seesaw mechanism for Higgs doublets and the scale of new physics, \emph{Phys. Lett. B.}, {\bf 674}, 2, 13 April 2009, Pages 117-121.
\bibitem{LSND}
  A.~Aguilar-Arevalo \emph{et al.} Evidence for neutrino oscillations from the observation of anti-neutrino(electron) appearance in a anti-neutrino(muon) beam. \emph{Phys. Rev. D.}, {\bf 64} 112007, 2001.
\bibitem{mini1}
  A.~Aguilar-Arevalo \emph{et al.} Event Excess in the MiniBooNE Search for $\nu_\mu \rightarrow \nu_e$ Oscillations. \emph{Phys. Rev. Lett.}, {\bf 105} 181801, 2010.
\bibitem{mini2}
  A.~Aguilar-Arevalo \emph{et al.} Improved Search $\nu_\mu \rightarrow \nu_e$ Oscillations in the MiniBooNE Experiment. \emph{Phys. Rev. Lett.}, {\bf 110} 161801, 2013.
\bibitem{reactor_anom1}
  G. Mention, M. Fechner, Th. Lasserre, Th.A. Mueller, D. Lhuillier, \emph{et al.} The Reactor Antineutrino Anomaly. \emph{Phys. Rev. D.}, {\bf 83} 073006, 2011.
\bibitem{reactor_anom2}
  Th. A. Mueller, D. Lhuillier, M. Fallot, A. Letourneau, S. Cormon, \emph{et al.} Improved Predictions of Reactor Antineutrino Spectra. \emph{Phys. Rev. C.}, {\bf 83} 054615, 2011.
\bibitem{gallex1}
  P.~Anselmann~\emph{et al.} First results from the Cr-51 neutrino source experiment with the GALLEX detector. \emph{Phys. Lett. B.}, {\bf 342} 440-450, 1995
\bibitem{gallex2}
  W.~Hampel~\emph{et al.} Final results of the Cr-51 neutrino source experiments in GALLEX. \emph{Phys. Lett. B.}, {\bf 420} 114-126, 1998.
\bibitem{sage1}
  Dzh.~N.~Abdurashitov, V.N.~Gavrin, S.V. Girin, V.V. Gorbachev, Tatiana V. Ibragimova, \emph{et al.} The Russian-American gallium experiment (SAGE) Cr neutrino source measurement. \emph{Phys. Rev. Lett.}, {\bf 77} 4708-4711, 1996.
\bibitem{sage2}
  J.~N.~Abdurashitov~\emph{et al.} Measurement of the response of the Russian-American gallium experiment to neutrinos from a Cr-51 source. \emph{Phys. Rev. C.}, {\bf 59} 2246-2263, 1999.
\bibitem{MINOS}
  Alexandre B. Sousa (for the MINOS and MINOS+ Collaborations) First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}
\bibitem{SuperK}
  K. Abe~\emph{et al.} (Super-Kamiokande Collaboration) Limits on sterile neutrino mixing using atmospheric neutrinos in Super-Kamiokande. \emph{Phys. Rev. D}, {\bf 91}, 052019, 2015, \url{http://dx.doi.org/10.1103/PhysRevD.91.052019}
\bibitem{DayaBay}
  F. P. An~\emph{et al.} (Daya Bay Collaboration) Search for a Light Sterile Neutrino at Daya Bay. \emph{Phys. Rev. Lett.} {\bf 113}, 141802, 2014, \url{http://dx.doi.org/10.1103/PhysRevLett.113.141802}
